Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-22 WSSSPE2_report/WSSSPE2_review.tex
Section 7: More minor changes.

The discussion of incentivizing those in science acknowledged that the
publications system was far from perfect for incentivizing good software work.
Nonetheless discussion focused on bringing software people into publications.
There were two main suggestions. The first is to focus on end users of software
and encourage them to cite software the ``right" way. James Howison suggested
that his research showed that few projects were making a formal request for
citation (but that authors weren't necessarily following those suggestions
anyway) \cite{howison2015jasist} he suggested making access to the software
conditional on a license that requires citation. Others found this ``too
confrontational" and preferred to concentrate on making it easier to do the
right thing. The second was focusing on those leading software projects, and the
group was more enthusiastic about ``forcing" PIs to include their ``software
people" on publications, although there were few ideas on how exactly to do
this. Another technique mentioned was that when scientific software projects are
hosted in organizations like Apache, the scientific contributors can benefit
from building their reputations, perhaps yielding job offers that they can use
to negotiate better job and career packages.

The discussion on incentivizing those outside science focused on accessing the
well of affection that those working in software have for scientific research.
How can the interest and skills of this group be marshaled towards sustainable
contributions? The is evidence that the migration of scientific software
projects to the Apache Foundation has created opportunities for those not
employed in the scientific center to contribute to projects initiated by
scientists (especially where there is cross-over with industry needs, such as
provenance and workflow).

@@

In the remaining discussion sessions, the group chose to split into three
subgroups to discuss a version of these topics: {\em citation ecosystems}, {\em
taxonomy of contributorship/guidelines for software for tenure review}, and {\em
prizes}. The subgroups were asked to clearly identify
\begin{itemize}
\item the problem to be solved
\item steps towards a solution
\end{itemize}

@@

The following general points were initially discussed: 
\begin{itemize}
\item The FLASH project was suggested as an example of how something 
like this has been done.
\item This data could be a file that can be associated with a citation to the 
software, either through use of a DOI for the credit file, or by uploading the 
credit files as associated with the paper.
\item This idea could also be applied to data
\item This data (the credit file) should be part of the metadata that is outside
of paywalls, like citations are now.
\item It was suggested that there should also be a separate file to track
software dependencies, but this is not the same file. 
\end{itemize}

The group came up with the following actions to be performed:
\begin{enumerate}
\item Build a tool that can automatically determine who the contributors are 
(from a git or other repo), then allows the user to manually edit the output to 
add/remove people, define roles. 
\item Work with repositories to encourage them to provide the information we 
need based on what they already store.
\item Define what a citation file should look like and what it should be called.
\item Test adoption, for example, with LBNL.
\item Create credit file for a set of software.
\item Build a validator (and perhaps a visualizer) for credit files.
\item Write a tool to collect files and visualize/output interconnections (which 
software is used with which), based on an existing project.
\item When we (the group members) write papers, we should track the software 
we use, and encourage the software developers to make their software citable 
and create credit files.
\item Build a tool to export the credit file to BibTex and other citation styles.
\item Make sure the BibTex entries (exported from internal data) are somewhat 
standardized so that they can be imported into papers.  Also make sure that 
standard LaTex style files understand and accept these entries.
\end{enumerate}

>>>>

The discussion of incentivizing those in science acknowledged that the
publications system was far from perfect for incentivizing good software work.
Nonetheless discussion focused on bringing software people into publications.
There were two main suggestions. The first is to focus on end users of software
and encourage them to cite software the ``right" way. James Howison suggested
that his research showed that few projects were making a formal request for
citation (but that authors were not necessarily following those suggestions
anyway) \cite{howison2015jasist}. He suggested making access to the software
conditional on a license that requires citation. Others found this ``too
confrontational" and preferred to concentrate on making it easier to do the
right thing. The second was focusing on those leading software projects, and the
group was more enthusiastic about ``forcing" PIs to include their ``software
people" on publications, although there were few ideas on how exactly to do
this. Another technique mentioned was that when scientific software projects are
hosted in organizations like Apache, the scientific contributors can benefit
from building their reputations, perhaps yielding job offers that they can use
to negotiate better job and career packages.

The discussion on incentivizing those outside science focused on accessing the
well of affection that those working in software have for scientific research.
How can the interest and skills of this group be marshaled towards sustainable
contributions? There is evidence that the migration of scientific software
projects to the Apache Foundation has created opportunities for those not
employed in the scientific center to contribute to projects initiated by
scientists (especially where there is cross-over with industry needs, such as
provenance and workflow).

@@

In the remaining discussion sessions, the group chose to split into three
subgroups to discuss a version of these topics: {\em citation ecosystems}, {\em
taxonomy of contributorship/guidelines for software for tenure review}, and {\em
prizes}. The subgroups were asked to clearly identify
\begin{itemize}
\item the problem to be solved
\item steps towards a solution.
\end{itemize}

@@

The following general points were initially discussed: 
\begin{itemize}
\item The FLASH \choinote{cite} project was suggested as an example of how something 
like this has been done.
\item This data could be a file that can be associated with a citation to the 
software, either through use of a DOI for the credit file, or by uploading the 
credit files as associated with the paper.
\item This idea could also be applied to data
\item This data (the credit file) should be part of the metadata that is outside
of paywalls, like citations are now.
\item It was suggested that there should also be a separate file to track
software dependencies, but this is not the same file. 
\end{itemize}

The group came up with the following actions to be performed:
\begin{enumerate}
\item Build a tool that can automatically determine who the contributors are 
(from a git or other repo), then allows the user to manually edit the output to 
add/remove people, define roles. 
\item Work with repositories to encourage them to provide the information we 
need based on what they already store.
\item Define what a citation file should look like and what it should be called.
\item Test adoption, for example, with LBNL.
\item Create credit file for a set of software.
\item Build a validator (and perhaps a visualizer) for credit files.
\item Write a tool to collect files and visualize/output interconnections (which 
software is used with which), based on an existing project.
\item When we (the group members) write papers, we should track the software 
we use, and encourage the software developers to make their software citable 
and create credit files.
\item Build a tool to export the credit file to BibTex and other citation styles.
\item Make sure the BibTex entries (exported from internal data) are somewhat 
standardized so that they can be imported into papers.  Also make sure that 
standard \LaTeX style files understand and accept these entries.
\end{enumerate}
